
---
output:
  html_document:
    df_print: kable
---


```{r echo=FALSE, message=FALSE, warning=FALSE}
# QUESTION: Which NYC airport has the longest delays for a given day of the week?

# If you do nto have one or more of the packages used below, please install them
# by connecting to the internet, opening an R session and running this command (without the #):
# install.packages(c("nycflights13", "dplyr", "ggplot2", "lubridate"))
library(nycflights13) # data
library(dplyr)        # cleaning functions
library(ggplot2)      # plotting functions
library(lubridate)    # date-time processing
library(kableExtra)

# Make a new data frame named delays. Make it like this:
#   take flights and then
#   mutate it to create a new date variable
#     and a new weekday variable, then
#   select just the origin, date, weekday, and dep_delay columns, then
#   filter out the rows where dep_delay equals NA
delays <-
  flights %>%
  mutate(date = ymd(paste(year, month, day)), 
         weekday = wday(date, label = TRUE, abbr = FALSE)) %>% 
  select(origin, date, weekday, dep_delay) %>%
  filter(!is.na(dep_delay)) 

# MOTIVATION: Delays fluctuate throughout the year

# Make a new data frame named year. Make it like this:
#   take delays and then
#   group its rows into unique combinations of origin and date, then 
#   summarise the data by calculating the mean dep_delay for each group
year <-
  delays %>% 
  group_by(origin, date) %>% 
  summarise(mean_delay = mean(dep_delay))

# METHOD: Choose a day of the week and aggregate delays by airport

# Pick a day of the week
dow <- "Saturday"

# Make a new data frame named weekday. Make it like this:
#   take delays and then
#   filter out rows where weekday does not equal the day above, then
#   group its rows by origin, then 
#   summarise the data by calculating the mean dep_delay for each group
weekday <- 
  delays %>%
  filter(weekday == dow) %>%
  group_by(origin) %>%
  summarise(mean_delay = mean(dep_delay))
```



## Goal

Which airport near New York City has the shortest departure delays on Saturday?

## Motivation

Airline passengers have choices when they choose to fly in to or out of New York City. Three major airports serve the area:

* **JFK** - John F Kennedy International Airport
* **LGA** - LaGuardia Airport
* **EWR** - Newark Liberty International Airport

How should passengers choose which airport to depart from? This study compares the airports on an obvious criteria: _departure delays_, the amount of time that passes between when a plane is scheduled to depart and when it actually departs. The amount of delay varies by ariport and fluctuates throughout the year, as seen in the graph below.

```{r echo=FALSE, message=FALSE}
# Plot the mean departure delay over time by airport
ggplot(year, aes(x = date, y = mean_delay, color = origin)) + 
  geom_point(alpha = 0.2) +
  geom_smooth(se = FALSE) +
  ggtitle("Smoothed daily mean delays") + 
  ylab("Mean delay (m)") + 
  theme_bw()
```


Since passengers can normally plan which day of the week they travel on, we will condition our analysis on a single day of the week: Saturday.

## Data

We will examine all flights that originated from a New York City airport in 2013. The data is provided by the Bureau of Transportation Statistics (BTS) and is available for download from <http://www.transtats.bts.gov/DL_SelectFields.asp?Table_ID=236>.

## Conclusion

```{r echo=FALSE, message=FALSE}
# Plot the mean delay by airport for the selected day of the week
ggplot(weekday, aes(x = origin)) + 
  geom_bar(aes(weight = mean_delay)) +
  ggtitle(paste("Expected", dow, "departure delay", sep = " ")) + 
  ylab("Mean delay (m)") +
  xlab("")



```

   
As the plot above shows, `r c("EWR" = "Newark", "JFK" = "JFK", "LGA" = "LaGuardia")[[weekday$origin[which.min(weekday$mean_delay)]]]` had the shortest mean departure delay on `r dow`, with a mean delay of `r round(min(weekday$mean_delay), 2)` minutes.

The complete results are in the table below.

```{r results="asis", echo=FALSE}
kable(weekday) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```




